
---
title: "A Quick introudction for chaolab utility"
output: github_document
---


<!-- README.md is generated from README.Rmd. Please edit that file -->


```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "", 
                      fig.retina=2,
                      fig.align='center',
                      fig.width = 7, fig.height = 5,
                      warning = FALSE, message = FALSE)
```

#overview 

```{r setup}
library(chaoUtility)
```


`chaoUtility` focuses on some useful tools for chao lab, include:

general functions:[`Boot_p()`](#boot_p) and `checktype()`

phylogeny fucntions:

-[`phylo2phytree()`](#phylo2phytree) : input phylo object([ultrmetric](#examples-ultrametictree) or [non-ultrametric](#examples-non-ultrametictree)), return chaophytree object 




-[`phyExpandData()`](#phyexpanddata) : input abundance data, label,chaophytree object, return tibble with abundance or [incidence data](#examples-incidence-simple-data).



-[`phyExpandData()`](#phyexpanddata) : input abundance data, label,chaophytree object, return tibble with abundance or [incidence data](#examples-incidence-simple-data).


-[`phylengthbyT()`](#phylengthbyt) : input vector of ageT, chaophytree object, return matrix with label and new branch.length (default `rootExtend =T` ,if `rootExtend=T and ageT>treeH `, root.length=ageT-treeH), [non ultrametric tree example](#examples-non-ultrametric-tree-by-reference-t)

-[`phyBranchAL()`](#phybranchal) : input abundance data,phylo object,vector of ageTs,  return Chaophyabu object  (default `rootExtend =T` ,`removeAbu0=T`)



-`phylo2chaolabphy()`



## HOW TO RUN chaoUtility:
The `chaoUtility` package can be downloaded with a standard R installation procedure using the following commands.

```{r eval=FALSE}
## install the latest version from github
install.packages('devtools')
library(devtools)
install_github('chaolab2019/chaoUtility')


```


## Boot_p 



We first describe the main function `Boot_p()` with default arguments: 

```{r eval=FALSE}
Boot_p(x,zero=TRUE,Bootype="One",datatype="abundance")
```
<table style="width:100%;">
<colgroup>
<col width="20%">
<col width="80%">
</colgroup>
<thead>
<tr class="header">
<th align="center">Argument</th>
<th align="left">Description</th>
</tr>
</thead>
<tbody>

<tr class="odd">
<td align="center"><code>x</code></td>
<td align="left">a <code>matrix</code>, <code>data.frame</code>, <code>lists</code> of species abundances or incidence data.</td>
</tr>
<tr class="even">
<td align="center"><code>zero</code></td>
<td align="left">reserves zero frequency or not. Default is TRUE.</td>
</tr>
<tr class="odd">
<td align="center"><code>Bootype</code></td>
<td align="left"> <code>character</code> : "One" or "JADE" or "SAR". `Bootype = "One"` or `Bootype = "JADE"` or `Bootype = "SAR"`<br> "One": from #99 2013 Entropy and the species accumulation curve Appendix S2: Estimating species relative abundance. <br>"JADE": from #107 Unveiling the species-rank abundance distribution by generalizing the Good-Turing sample coverage theory. <br>"SAR": :sampleing which was drawn without replacement (based on one parameter estimated method)</td>
<tr class="even">
<td align="center"><code>datatype</code></td>
<td align="left">data type of input data: individual-based abundance data (`datatype = "abundance"`), or species by sampling-units incidence frequency (`datatype = "incidence_freq"` or `datatype = "incidence"`),or species by sampling-units incidence matrix (`datatype = "incidence_raw"`).</td>

</tbody>
</table>

## EXAMPLES


without replacement:
```{r }
library(chaoUtility)
data(bird)
bird.abun<- bird$abun
Boot_p(bird$abun,zero=FALSE,Bootype="SAR",datatype="abundance",rho=0.8)

data(incdata)
Boot_p(incdata,zero=FALSE,Bootype="SAR",datatype="incidence",rho=0.8)
```


with replacement:One
```{r }
library(chaoUtility)
data(bird)
bird.inc <- bird$inci
bird.abun<- bird$abun
Boot_p(bird$abun, datatype="abundance")
```
## References
- Chao, A., Wang, Y. T., and Jost, L. (2013). Entropy and the species accumulation curve: a novel estimator of entropy via discovery rates of new species. Methods in Ecology and Evolution, 4, 1091-1110. 
- Chao, A., Hsieh, T. C., Chazdon, R. L., Colwell, R. K., and Gotelli, N. J. (2015). Unveiling the species-rank abundance distribution by generalizing the Good-Turing sample coverage theory. Ecology 96, 1189-1201.

[back](#overview)



## phylo2phytree


### examples ultrametictree

```{r }
library(chaoUtility)
data(treesample)
newphy<-phylo2phytree(treesample)

class(newphy)
```


```{r }
newphy$tips

newphy$nodes

newphy$phytree


newphy$treeH
```

```{r }
library(ape)
plot(treesample)
tiplabels()
nodelabels()
library(dplyr)
nodetext<-newphy$phytree %>% filter(tgroup!="Tip") %>% pull(label)
nodelabels(text=nodetext,adj=c(0,2.2))
edgelabels(treesample$edge.length, bg="black", col="white", font=2)
```
[back](#overview)


### examples non ultrametictree

```{r }
text<-"(((((((A:4,B:4):6,C:5):8,D:6):3,E:21):10,((F:4,G:12):14,H:8):13):13,((I:5,J:2):30,(K:11,L:11):2):17):4,M:56);"
library(ape)
tree2<-read.tree(text=text)

library(chaoUtility)
tree2.phytree<-phylo2phytree(tree2)

library(dplyr)
nodetext<-tree2.phytree$phytree %>% filter(tgroup!="Tip") %>% pull(label)
plot(tree2)
nodelabels(text=nodetext)
edgelabels(tree2$edge.length, bg="black", col="white", font=2)

tree2.phytree$treeH

as.data.frame(tree2.phytree$phytree)

tree2.phytree$leaves

tree2.phytree$nodes


```
[back](#overview)


### examples non ultrametictree

```{r }
text<-"(((((((A:4,B:4):6,C:5):8,D:6):3,E:21):10,((F:4,G:12):14,H:8):13):13,((I:5,J:2):30,(K:11,L:11):2):17):4,M:56);"
library(ape)
tree2<-read.tree(text=text)

library(chaoUtility)
tree2.phytree<-phylo2phytree(tree2)

library(dplyr)
nodetext<-tree2.phytree$phytree %>% filter(tgroup!="Tip") %>% pull(label)
plot(tree2)
nodelabels(text=nodetext)
edgelabels(tree2$edge.length, bg="black", col="white", font=2)

tree2.phytree$treeH

as.data.frame(tree2.phytree$phytree)

tree2.phytree$leaves

tree2.phytree$nodes


```


## phyExpandData


### EXAMPLES:abundance data

```{r }
library(chaoUtility)
data(phybird.new)
bird.abu <- phybird.new$abun
bird.lab <- rownames(phybird.new$abun)
bird.phy <- phybird.new$chaophytree

bird.abu

bird.lab


phyExpandData(x=bird.abu, labels=bird.lab, phy=bird.phy, datatype="abundance")
```



### examples incidence simple data
```{r }
library(chaoUtility)
data(phyincisimple)
data.inc <- phyincisimple$inci.simple.data
data.lab<-rownames(data.inc)
phy.inc<-phyincisimple$tree.simple.phytree
```


[back](#overview)

### examples incidence simple data
```{r }
library(chaoUtility)
data(phyincisimple.new)
data.inc <- phyincisimple.new$inci.simple.data
data.lab<-rownames(data.inc)
phy.inc<-phyincisimple.new$tree.simple.phytree



phylotree<-phyincisimple.new$tree.simple
plot(phylotree)
nodetext<-phy.inc$phytree %>% filter(tgroup!="Tip") %>% pull(label)
nodelabels(text=nodetext)
edgelabels(phylotree$edge.length, bg="black", col="white", font=2)
```

```{r }
data.inc

dataNtree<-phyExpandData(data.inc, labels=data.lab, phy=phy.inc, datatype="incidence_raw")
as.data.frame(dataNtree)
```




[back](#overview)




## phyLengthbyT


### EXAMPLES

```{r }
library(chaoUtility)
data(phybird)
bird.phy <- phybird$chaophytree


bird.phy$treeH

phyLengthbyT(Ts=c(90,75,55), phy=bird.phy, datatype="abundance")
```


[back](#overview)


### examples non ultrametric tree by reference t
```{r }
text<-"(((((((A:4,B:4):6,C:5):8,D:6):3,E:21):10,((F:4,G:12):14,H:8):13):13,((I:5,J:2):30,(K:11,L:11):2):17):4,M:56);"
library(ape)
tree2<-read.tree(text=text)
```

### examples non ultrametric tree by reference t
```{r }
text<-"(((((((A:4,B:4):6,C:5):8,D:6):3,E:21):10,((F:4,G:12):14,H:8):13):13,((I:5,J:2):30,(K:11,L:11):2):17):4,M:56);"
library(ape)
tree2<-read.tree(text=text)

library(chaoUtility)
phytree<-phylo2phytree(tree2)

phytree$treeH

phyLengthbyT(Ts=c(75,55,50), phy=phytree, datatype="abundance",rootExtend=T)

library(dplyr)
nodetext<-phytree$phytree %>% filter(tgroup!="Tip") %>% pull(label)
plot(tree2)
nodelabels(text=nodetext)
edgelabels(tree2$edge.length, bg="black", col="white", font=2)

as.data.frame(phytree$phytree)
```


[back](#overview)


## #phybranchal


### EXAMPLES

```{r }
 data(AbuALdata)
 adata<-AbuALdata$abudata
 atree<-AbuALdata$tree
 vdata<-adata$EM
 names(vdata)<-rownames(adata)
 vdata
 
 refTs<-c(400,250,100)
 result<-phy_BranchAL_Abu(atree,vdata,datatype="abundance",refTs)
 
 ##final branch.abu:removed abu=0
 treeNabu<-result$treeNabu
 treeNabu %>% print(n = Inf)
 
 ##final treeH:removed abu=0
 result$treeH
 
 ##final branch.length:removed abu=0
 result$BLbyT
 
 ####result$treeNabu is an object of tbl_tree could change to phylo
 tlb2phylo<-as.phylo(treeNabu)
 plot(tlb2phylo)
 
 
 ###this is the original tree
 plot(atree)
 
```

[back](#overview)
