
---
title: "A Quick introudction for chaolab utility"
output: github_document
---


<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "", 
                      fig.retina=2,
                      fig.align='center',
                      fig.width = 7, fig.height = 5,
                      warning = FALSE, message = FALSE)
```

```{r setup}
library(chaoUtility)
```
`chaoUtility` focuses on some useful tools for chao lab, include bootstrap and checktype functions

## HOW TO RUN chaoUtility:
The `chaoUtility` package can be downloaded with a standard R installation procedure using the following commands.

```{r eval=FALSE}
## install the latest version from github
install.packages('devtools')
library(devtools)
install_github('chaolab2019/chaoUtility')

## import packages

```
## MAIN FUNCTION: Boot_p()

We first describe the main function `Boot_p()` with default arguments: 

```{r eval=FALSE}
Boot_p(x,zero=TRUE,Bootype="One",datatype="abundance")
```
<table style="width:100%;">
<colgroup>
<col width="20%">
<col width="80%">
</colgroup>
<thead>
<tr class="header">
<th align="center">Argument</th>
<th align="left">Description</th>
</tr>
</thead>
<tbody>

<tr class="odd">
<td align="center"><code>x</code></td>
<td align="left">a <code>matrix</code>, <code>data.frame</code>, <code>lists</code> of species abundances or incidence data.</td>
</tr>
<tr class="even">
<td align="center"><code>zero</code></td>
<td align="left">reserves zero frequency or not. Default is TRUE.</td>
</tr>
<tr class="odd">
<td align="center"><code>Bootype</code></td>
<td align="left"> <code>character</code> : "One" or "JADE". `Bootype = "One"` or `Bootype = "JADE"` <br> "One": from #99 2013 Entropy and the species accumulation curve Appendix S2: Estimating species relative abundance. <br>"JADE": from #107 Unveiling the species-rank abundance distribution by generalizing the Good-Turing sample coverage theory. </td>
<tr class="even">
<td align="center"><code>datatype</code></td>
<td align="left">data type of input data: individual-based abundance data (`datatype = "abundance"`), or species by sampling-units incidence frequency (`datatype = "incidence_freq"` or `datatype = "incidence"`),or species by sampling-units incidence matrix (`datatype = "incidence_raw"`).</td>

</tbody>
</table>
## EXAMPLES

```{r }
library(chaoUtility)
data(bird)
bird.inc <- bird$inci
bird.abun<- bird$abun
Boot_p(bird$abun, datatype="abundance")
```
## References
- Chao, A., Wang, Y. T., and Jost, L. (2013). Entropy and the species accumulation curve: a novel estimator of entropy via discovery rates of new species. Methods in Ecology and Evolution, 4, 1091-1110. 
- Chao, A., Hsieh, T. C., Chazdon, R. L., Colwell, R. K., and Gotelli, N. J. (2015). Unveiling the species-rank abundance distribution by generalizing the Good-Turing sample coverage theory. Ecology 96, 1189-1201.
